---
title: "FES 542 Lab 1"
author: "Kenneth Loonam"
date: "10/1/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
require(tidyverse)
```

## 1

```{r}
q_1_data <- tibble(
  meters = c((0:10)*10),
  ins_a = c(40,30,40,25,10,10,15,20,15,35,25),
  ins_b = c(2,3,0,4,29,22,18,8,10,1,2)
)

lm_fine <- with(q_1_data, lm(ins_b ~ ins_a))

ggplot(data = q_1_data, aes(x = ins_a, y = ins_b)) +
  geom_point() +
  theme_classic() +
  labs(x = "Insect A", y = "Insect B", title = "Correlation of Insect Abundance at Fine Scale") + 
  geom_abline(slope = -0.7575, intercept = 27.2491)
```

## 2

```{r}
summary(lm_fine)
```


## 3

```{r}
q_3_data <- tibble(
  meters = c((0:10) * 2000),
  ins_a = c(8000,4800,2000,6000,0,400,400,10800,4000,2000,8000),
  ins_b = c(16000,8000,400,12000,400,400,3200,16000,6000,3600,8800)
)

lm_broad <- with(q_3_data, lm(ins_b ~ ins_a))

ggplot(data = q_3_data, aes(x = ins_a, y = ins_b)) +
  geom_point() +
  theme_classic() +
  labs(x = "Insect A", y = "Insect B", 
       title = "Correlation of Insect Abundance at Broad Scale") + 
  geom_abline(slope = 1.508, intercept = 437.475)

summary(lm_broad)
```

## 4

At fine scales, insects A and B are negatively correlated, however, at broad 
scales (e.g. larger extent and grain), insects A and B are positively 
correlated. The negative correlation at fine scales supports two hypotheses: 
that insects A and B compete or that insects A and B select for different 
habitats or resources at fine scales. The positive correlation at broad scales 
also supports multiple hypotheses. A and B could utilize similar habitats at 
broad scales. They could also be mutualists, which is unlikely given the fine 
scale observations. The observations at both scales suggest that A and B either 
compete but use similar habitat or that they select habitat similarly at broad 
scales but not at fine scales.

## 5

The mechanisms driving abundance of the two species might be different at broad
and find scales (as detailed in the hypotheses in part 4). It's also plausible 
that the relationships are spurious, particularly the fine scale observations. 
Measuring abundances in 100 sq cm plots over and extent of 100 m could be
subject to extreme variance. If a similar survey were repeated along different
transects, it would lead to more certainty.

## 6

The grain (0.1 to 10 sq m) and the extent (100 to 20000 m) vary between surveys.

## 7

In Starkey, the large mammals are hypothesized to have multiple relationships.
Elk and mule deer could select for similar habitats or competitively exclude each
other. Similarly, coyotes could use different habitat than cougars, actively
avoid cougars, or seek out cougar kills to scavenge. Understanding the scale 
that these behaviors happen at is critical to teasing out the relationships. If 
mule deer select for similar habitat as elk, there might be a lot of overlap at 
the home range scale. However, they could still avoid elk within their home 
range, leading to different results when mule deer and elk abundance are
measured at different scales.